% ************************** Thesis Abstract *****************************
% Use `abstract' as an option in the document class to print only the titlepage and the abstract.
\begin{abstract2}

The following work has been published as part of the research described in this thesis.

 \section*{Papers}
  \begin{enumerate}
  \item Wallace, Duncan, and Tahar Kechadi. "Prediction of Frequent Out-Of-Hours’ Medical Use." ECML PKDD 2019 Workshops: Nemesis 2019, UrbReas 2019, SoGood 2019, IWAISe 2019, and Green Data Mining 2019,  Würzburg, Germany, September 16-20, 2019, Proceedings.
  \item Wallace, Duncan, and Tahar Kechadi. "Outlier Detection in Health Record Free-Text Using Deep Learning" 2019 41st Annual International Conference of the IEEE Engineering in Medicine and Biology Society (EMBC). Berlin, Germany, July 23-27, 2019.
   \item Wallace, Duncan, and Tahar Kechadi. "Analysis of EHR Free-text Data with Supervised Deep Neural Networks." CSCE'18: The 2018 World Congress in Computer Science, Computer Engineering \& Applied Computing, Las Vegas, Nevada, USA, 30 July-02 August 2018.
  \item Wallace, Duncan, and Tahar Kechadi. "Abbreviation and Acronym Identification and Expansion Within Medical Health Records." The 9th International Conference on e-Health, Lisbon, Portugal, July 20-22. IADIS, 2017.
  \item  Wallace, Duncan, and Tahar Kechadi. "Retrieval and Clustering of Medicines Within Healthcare Data Records." Proceedings of the International Conference on Big Data and Advanced Wireless Technologies. ACM, November 10-11,  2016.
  \end{enumerate}
  \section*{Posters}
  \begin{enumerate}
  \item Wallace, Duncan, and Tahar Kechadi. "Frequent attender Classification through Electronic Health Record Analysis" Science and Information Computing Conference 2019, London, United Kingdom. 16-17 July, 2019. 
  \end{enumerate}
  
%  @inproceedings{ibrahim2018optimal,
%  title={Optimal Moore Neighborhood Approach of Cellular Automaton Based Pedestrian Movement: A Case Study on the Closed Area},
%  author={Ibrahim, Najihah and Hassan, Fadratul Hafinaz},
%  booktitle={Science and Information Conference},
%  pages={57--71},
%  year={2018},
%  organization={Springer}
%}
  
 
% @inproceedings{paudice2018label,
%  title={Label sanitization against label flipping poisoning attacks},
%  author={Paudice, Andrea and Mu{\~n}oz-Gonz{\'a}lez, Luis and Lupu, Emil C},
%  booktitle={Joint European Conference on Machine Learning and Knowledge Discovery in Databases},
%  pages={5--15},
%  year={2018},
%  organization={Springer}
%}

%@inproceedings{huang2018roast,
%  title={ROAST: an open-source, fully-automated, realistic volumetric-approach-based simulator for TES},
%  author={Huang, Yu and Datta, Abhishek and Bikson, Marom and Parra, Lucas C},
%  booktitle={2018 40th Annual International Conference of the IEEE Engineering in Medicine and Biology %Society (EMBC)},
%  pages={3072--3075},
%  year={2018},
%  organization={IEEE}
%} 


%@inproceedings{wallace2019outlier,
% title={Outlier Detection in Health Record Free-Text Using Deep Learning},
%  booktitle={2019 41st Annual International Conference of the IEEE Engineering in Medicine and Biology Society (EMBC)},
%  pages={3072--3075},
%  year={2019},
%  organization={IEEE}
%}


\end{abstract2}
